% Appendix A
\setstretch{1}
\chapter{screenPerfect Installation Guide} % Main appendix title

\label{AppendixA} % Change X to a consecutive letter; for referencing this appendix elsewhere, use \ref{AppendixX}

\section{screenPerfect Code and Documentation}
The code of screenPerfect is located on the DVD included in this volume, and at \url{http://www.github.com/pretentiousgit/screenperfect-dev}. This DVD includes five games produced for the screenPerfect Engine: PornGame, psXXYborg, OM, Grimoire, and Cyborg Goddess.

\section{Device List}

\begin{enumerate}
\item{Large powerbar with surge protection and power supply cable}
\item{Server with Node.JS installed}
\item{Projector or monitor}
\item{Portable wireless hotspot}
\item{Touchscreen device with browser software}
\item{Optional: Speakers with subwoofer}
\item{keyboard}
\item{mouse}
\item{power supply}
\item{hdmi or vga cables as appropriate}
\end{enumerate}

\section{Software List}
\begin{enumerate}
\item{Google Chrome (preferred), on all devices likely to play.}
\item{screenPerfect will also run on Safari, but there are issues due to Google’s refusal to support the MP4/h.264 video codec (large, but good quality) and Apple’s similar refusal to support V8/webM (small, better, owned by Google).}
\item{NodeJS installed on screenPerfect’s host computer.}
\item{git installed on screenPerfect’s host computer.}
\item{Miro VideoConverter on a mac system}
\end{enumerate}

\section{Installation and Site Construction}

\subsection{Before Leaving For Site}
\begin{enumerate}
\item{Ensure NodeJS is installed and functional on main server.}
\item{Install latest repository of screenPerfect from GitHub.}
\item{Ensure all video files are present and accounted for in all appropriate formats.}
	\begin{enumerate}
		\item{webM}
		\item{MP4}
	\end{enumerate}
\item{Find surface to project on.}
\item{Lay out power cable and power bar.}
	\begin{enumerate}
		\item{power to wireless hotspot}
		\item{power to main server}
		\item{power to projector}
		\item{power to support tablet interface}
		\item{power to speakers}
		\item{monitor/projector}
		\item{keyboard}
		\item{mouse}
	\end{enumerate}
\item{Connect devices to their relevant power supplies.}
\item{Plug in green jack of speakers to headphone jack of main server.}
\item{Turn on all devices.}
\item{Turn off all mobile data for wireless hotspot device. If necessary, remove SIM.}
\item{Turn on wireless hotspot.}
\item{Turn on your main server.}
\item{Connect to your wireless hotspot.}
\item{Open Terminal (Mac), navigate to screenPerfect’s home directory, turn on screenPerfect}
	\begin{enumerate}
	\item{Navigate to the game folder you wish to run by typing "cd" and then the directory path at the prompt.}
	\item{type node screenPerfect, press enter.}
	\item{You will see a message saying “screenPerfect listening on port 3003”.}
	\end{enumerate}
\item{Open a new tab in Terminal (command-T)}
\item{Type ifconfig to get the IP address of your wireless hotspot.}
\begin{enumerate}
\item{look for the value called "inet" - it will be a \url{192.168.x.x} address if you are attached to the hotspot, and (probably) \url{10.x.x.x} if you’re attached to a wider network.}
\end{enumerate}
\item{On the Server, Open Chrome and navigate to \url{http://0.0.0.0:3003/client}}
\item{Open Terminal, and in the tab where you typed “node screenPerfect”  there will be a stream of messages about a heatbeat.}
\item{There should also be a video playing in your browser now.}
\item{On the Tablet Device, open Chrome. Replace the below values with the ip address located in step 14a}
\begin{enumerate}
\item{\url{http://x.x.x.x:3003/control}}
\end{enumerate}
\item{If all goes well, the “Start” button will appear onscreen on your tablet.}
\item{Touch the start button, and the projected video will change.}
\end{enumerate}

\section{Troubleshooting}

\subsection{On Start, Google Chrome Cannot Find Control Screen}

\begin{enumerate}
	\item{Check that the Tablet is on the wireless network provided by the hotspot.}
	\item{If not, repair the hotspot by disabling it, reenabling it, and reconnecting both the server and the tablet to the same network.}
\end{enumerate}

\subsection{The Control Device is frozen, or the videos are not changing.} 

Just about anything else, really, including the above.
\begin{enumerate}
	\item{Open the Terminal.}
	\item{Check the heartbeat stream for an error such as “Abort Trap,” “Bus Trap,” or “stream error in pipe.”}
	\begin{enumerate}
	\item{All of these are normal. They are a result of someone tapping the screen quickly enough to overwhelm the server.}
	\end{enumerate}
	\item{type node screenPerfect and press enter to restart the application.}
	\item{Refresh the tablet browser and press the start graphic to make sure the video is being served.}
	\item{Refresh the server browser to ensure everything is passed correctly.}
\end{enumerate}

\section{Tidbits and Tech Notes}

\begin{itemize}
	\item{People with little technical experience will sometimes have trouble determining if they have actually touched/interacted with the app appropriately. Some may accidentally exit the application on the tablet, and may be unsure of how to re-enter the experience.}
	\item{screenPerfect is most rewarding when played in isolated concentration somewhere quiet. It is challenging to concentrate on the video enough to get full effect if people are distracted.} 
	\item{screenPerfect crashes approximately once per three hours of play time, more in high temperatures.}
	\item{screenPerfect games work best in Android, as Android supports the webM/V8 codec. Apple devices require mp4s, which are approximately 4x the size.}
	\item{screenPerfect was developed in JavaScript using ExpressJS for NodeJS. screenPerfect was developed on a MacBook Pro for use with Chrome, but will function almost as well with Safari.}
\end{itemize}