% Appendix Template

\chapter{Game Jam Documentation} % Main appendix title

\label{AppendixC} % Change X to a consecutive letter; for referencing this appendix elsewhere, use \ref{AppendixX}

\lhead{Appendix C. \emph{Game Jam Documentation}} % Change X to a consecutive letter; this is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
%	Initial Questions
%----------------------------------------------------------------------------------------

\section{Interview Files}
Interview sound files can be found on the Supporting Materials DVD within the "Game Jam Interviews" folder.

\section{Questions To Ask Game Jammers}


\begin{itemize} 
\item[\tiny{$\blacksquare$}] What were you expecting when you came to the jam?
\item[\tiny{$\blacksquare$}] What features did you immediately want in your software?
\item[\tiny{$\blacksquare$}] How has your group process worked throughout the week?
\item[\tiny{$\blacksquare$}] How is your group process going today?
\end{itemize}

\subsection{Games List}
\begin{itemize} 
\item[\tiny{$\blacksquare$}] Porn Game by Maxwell Lander
\item[\tiny{$\blacksquare$}] Grimoire by Katie Foster and Mikayla Carson
\item[\tiny{$\blacksquare$}] Kill Fuck Marry by 
\item[\tiny{$\blacksquare$}] Mind Safe by Dann Toliver and Robby
\item[\tiny{$\blacksquare$}] Glitch95 by Arielle, Rebecca and Bronwyn
\item[\tiny{$\blacksquare$}] Omm by Brittany and Diana
\item[\tiny{$\blacksquare$}] Cyborg Goddess by Cara and Kate McKnyte
\item[\tiny{$\blacksquare$}] Empty Puppet by Danielle Hopkins and Dawn
\end{itemize}

\subsection{Bug Discovery}
\begin{itemize} 
\item[\tiny{$\blacksquare$}] Room Zero must be the first room edited.
\item[\tiny{$\blacksquare$}] Room Order cannot be altered in a meaningful way - ID is hidden from users
\item[\tiny{$\blacksquare$}] WebM video is unplayable on Apple devices
\item[\tiny{$\blacksquare$}] H.264 video is slow to unplayable on non-Apple devices
\end{itemize}


%----------------------------------------------------------------------------------------
%	Raw Notes
%----------------------------------------------------------------------------------------
\subsection{Features Requested by Game Jammers}

\begin{itemize} 
\item[\tiny{$\blacksquare$}] Sound effects on control input - requested by Arielle
\item[\tiny{$\blacksquare$}] Timed hotspots which appear and disappear on specific video cues
\item[\tiny{$\blacksquare$}] A game tracer that tracks which choices players make, and records their games
\item[\tiny{$\blacksquare$}] Gesture controls - pinch, zoom, throw - on touchpoints.
\item[\tiny{$\blacksquare$}] Tree View to visualize how a game is laid out
\item[\tiny{$\blacksquare$}] Rooms cannot be deleted - delete and reorder rooms
\item[\tiny{$\blacksquare$}] Games cannot be deleted - delete and reorder games
\item[\tiny{$\blacksquare$}] Copy and paste room layouts so that one does not have to recreate grids - done.
\item[\tiny{$\blacksquare$}] Hotspots that can move around the room.
\end{itemize}

\subsection{Notes from committed jammers about screenPerfect}

For Arielle, the most engaged of the jammers, the idea of turning any touch device into a custom console controller, with custom buttons, is engaging. The more traditional the game developer, the harder a time they had with the idea that they'd be showcasing content with the narrowest help from the new tool. The filmmakers were very impressed with the ability to not touch a darn thing and have considerable success.
