% Appendix Template

\chapter{Appendix D: Installing screenPerfect} % Main appendix title

\label{AppendixD} % Change X to a consecutive letter; for referencing this appendix elsewhere, use \ref{AppendixX}

\lhead{Appendix D. \emph{Installing screenPerfect}} % Change X to a consecutive letter; this is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
%	List of Software
%----------------------------------------------------------------------------------------

\section{screenPerfect Software Dependencies}
\begin{itemize}
\item[\tiny{$\blacksquare$}] \textbf{Node.JS} - a software framework for writing javascript on the server
\item[\tiny{$\blacksquare$}] \textbf{MongoDB} - a lightweight database framework utilizing JSON for file storage
\item[\tiny{$\blacksquare$}] \textbf{Daimio} - a dataflow language that stores the game-states of screenPerfect
\item[\tiny{$\blacksquare$}] \textbf{Git/GitHub} - version control software used to store and pull most recent versions of software
\item[\tiny{$\blacksquare$}] \textbf{screenPerfect} - software for branching narratives in video and image
\end{itemize}


Detailed installation instructions for ScreenPerfect on PC

1. Install node.js
	http://nodejs.org/ 
2. Install and run mondoDB following the instructions found below:
	http://docs.mongodb.org/manual/tutorial/install-mongodb-on-windows/
3. Either download and install Git/GitHub or download a copy of the SP working directory from pretentiousgit/screenperfect-dev, and install in local code directory. This should be somewhere that you are comfortable to find using the command line.

\section{Running screenPerfect from a command line}
1. From the command prompt, CD to the directory where you placed your copy of the screenPerfect files from web (for example, cd documents\nodejs\screenperfect-dev).
2. Type node screenperfect at the prompt.
The prompt should return:
\begin{lstlisting}
screenPerfect is listening on port 8888
\end{lstlisting}

3. Open your browser and head to:
\begin{lstlisting}
http://\textit{localhost}:8888/control will bring up a control interface
http://\textit{localhost}:8888/client will bring up a client screen
http://\textit{localhost}:8888/edit will bring up an editor, which will be only some use without a desktop.
\end{lstlisting}

\section{Accessing screenPerfect from a remote client}
\subsection{On Mac}
1. On your server machine, open a Terminal window, and enter 'ifconfig' at prompt.
2. ifconfig returns many results. You are looking for one that says 'inet.' The first ip address after inet is your localhost's dynamic address on your current wiFi network. 
3. To find screenPerfect, enter this address - for example, (192.168.10.10) - then :8888. http://\textit{192.168.10.10}:8888 should bring up the main screenPerfect game list. 

\begin{lstlisting}
http://\textit{192.168.10.10}:8888/control will bring up a control interface
http://\textit{192.168.10.10}:8888/client will bring up a client screen
http://\textit{192.168.10.10}:8888/edit will bring up an editor, which will be only some use without a desktop.
\end{lstlisting}

\subsection{On Windows}
1. Open a CMD window, and enter ipconfig at the prompt.
2. Take the number next to IPv4, for example, \textit{192.168.10.10}.
3. Open your browser and enter http://\textit{192.168.10.10}:8888/ to access the main game trees.

\begin{lstlisting}
http://\textit{192.168.10.10}:8888/control will bring up a control interface
http://\textit{192.168.10.10}:8888/client will bring up a client screen
http://\textit{192.168.10.10}:8888/edit will bring up an editor, which will be only some use without a desktop.
\end{lstlisting}
